Synthesis of Rules for Network Intrusion Detectors

### 2019-10-22 paper/paper.tex
pass on technique

\begin{wrapfigure}[14]{r}{0.5\textwidth}
  \centering
  \vspace{-5ex}  
  \begin{tikzpicture}
    \begin{axis}[
        ybar stacked,
        enlargelimits=0.15,
        legend style={at={(0.5,-0.15)},
          anchor=north,legend columns=-1},
        ylabel={\#rules},
        symbolic x coords={gen, http, icmp, udp, tcp},
        xtick=data,
        nodes near coords,
        nodes near coords align={vertical},
      ]
      \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
      \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
      \legend{relevant, irrelevant}
    \end{axis}
  \end{tikzpicture}
  \caption{\label{fig:distribution-rules-protocol}Distribution of
    rules per protocol and relevance for triggering action.}
\end{wrapfigure}
It is important to note that options can be specific to a given
protocol or general and they can influence or not triggering an
action. Figure~\ref{fig:distribution-rules-protocol} shows the
distributions of these options per protocol and their relevance for
triggering rule actions. This plot indicates that the problem search
space is defined based on the protocol of the malicious traffic. For
instance, if \tname{} detects that the attack involves \CodeIn{http}
traffic, then 34 different rule options will be considered in the
creation of the rule.

\Fix{describe extraction...}

@@

\begin{figure*}[t]
\centering
\includegraphics[trim=10 360 182 130,clip,width=0.95\textwidth]{figs/nids-workflow}
\caption{\tname\ workflow.}
\label{fig:overview}
\end{figure*}

>>>>

\begin{wrapfigure}[14]{r}{0.5\textwidth}
  \centering
  \vspace{-5ex}  
  \begin{tikzpicture}
    \begin{axis}[
        ybar stacked,
        enlargelimits=0.15,
        legend style={at={(0.5,-0.15)},
          anchor=north,legend columns=-1},
        ylabel={\#rules},
        symbolic x coords={gen, http, icmp, udp, tcp},
        xtick=data,
        nodes near coords,
        every node near coord/.append style={font=\scriptsize},        
        nodes near coords align={vertical},
      ]
      \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
      \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
      \legend{relevant, irrelevant}
    \end{axis}
  \end{tikzpicture}
  \caption{\label{fig:distribution-rules-protocol}Distribution of
    rules per protocol and relevance for triggering action.}
\end{wrapfigure}
It is important to note that options can be specific to a given
protocol or general and they can influence or not triggering an
action. The purpose of rule options \CodeIn{msg}, for instance, is to
print on the output an informative message indicating what kind of
intrusion has been
observed. Figure~\ref{fig:distribution-rules-protocol} shows the
distributions of these options per protocol and their relevance for
triggering rule actions. This plot indicates that the problem search
space is defined based on the protocol of the malicious traffic. For
instance, if \tname{} detects that the attack involves \CodeIn{http}
traffic, then 34 different rule options will be considered in the
creation of the rule.

To infer options, \tname{} parses the protocol message, looking for
fields in the message associated with options. For example, fields
\Fix{...and...}  can be used to determine the option
\Fix{...}. Unfortunately, inferring options at the field granularity
level is insufficient to capture content options (see
Section~\ref{sec:example-suricata-rules}), which refer to parts of the
payload of the message and cannot be directly extracted directly from
the fields. Handling this case is critically important as most rules
contain content options. Figure~\ref{fig:distribution-contents} shows
the histogram of number of content options per rule (up to 10) for the
Suricata public rule database \Fix{cite}. Rules without content
options is an exception and half of the rules contain at least two
content options.
\begin{wrapfigure}[12]{r}{0.5\textwidth}
  \centering
  \vspace{-5ex}  
  \begin{tikzpicture}
    \begin{axis}[
        bar width=8pt,
        scaled ticks=false,
        tick label style={/pgf/number format/fixed},
        ybar stacked,
        enlargelimits=0.1,
        legend style={at={(0.5,-0.15)},
          anchor=north,legend columns=-1},
        symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
        xtick=data,
        nodes near coords,
        every node near coord/.append style={font=\tiny},
        nodes near coords align={vertical},
      ]
      \addplot coordinates {(0,25) (1,10222) (2,4956) (3,5589)
        (4,1651) (5,1457) (6, 729) (7, 471) (8, 313) (9, 137) (10, 126)};
    \end{axis}
  \end{tikzpicture}
  \caption{\label{fig:distribution-contents}Histogram of number of
    content options per rule (up to 10).}
\end{wrapfigure}
To find the values associated with these options, \tname{} splits the
payload of the malicious message in tokens using natural language
delimiters, such as $\backslash$t, $\backslash$n, $\backslash$r, and
spaces. The rationale for this decision is that we observed, by
inspecting existing rules, that content options typically refer to
sequence of characters in the payload of messages separated by these
delimiters. \Fix{describe extraction of options not in the
  message. give examples! search only makes sense if this exists!}

@@

%% \begin{figure*}[t]
%% \centering
%% \includegraphics[trim=10 360 182 130,clip,width=0.95\textwidth]{figs/nids-workflow}
%% \caption{\tname\ workflow.}
%% \label{fig:overview}
%% \end{figure*}
